Distributed Nonlinear Feedback Design with Separable Control-Contraction Metrics

### 2017-02-05 R0V0.tex
Change theorem statement and propagate changes

\begin{theorem}\label{thm:main result}
	If there exists a solution to the matrix inequalities
	\begin{subequations}
		\begin{equation}\label{eq:MI:SSCCM}
			-\dot{W}+A+WA^\top+BY+(BY)^\top+2\lambda W\prec0
		\end{equation}
		with $W=\mathbin{\mathtt{diag}}(W_1,\ldots,W_N)$,
		\begin{equation}\label{eq:Ws:SSCCM}
			W_1\succ0,\ldots,W_N\succ0
		\end{equation}
		and
		\begin{equation}\label{eq:YinXi}
			Y\in\Xi.
		\end{equation}
	\end{subequations}
	Then, $W$ is a sum-separable control-contraction metric for system \eqref{eq:general system} and there exists a solution to Problem~\ref{problem formulation}. In particular, the graph $\mathscr{G}_c$ describing the communication network among different controllers specifies the set $\Xi$.
\end{theorem}

\begin{proof}[Proof of Theorem \ref{thm:main result}]
	By assumption, the functions $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ are a solution to the MI \eqref{eq:MI formulation}. Apply the coordinate change $\eta=M\delta$ and define the matrices $W=M^{-1}$ and $K=YW^{-1}$. Since $W$ is diagonal, the structure of $Y$ is preserved and, consequently, $K\in\Xi$.

\mypara	The MI \eqref{eq:MI formulation} implies that the inequality
	\begin{equation*}
		\delta_x^\top\bigg(\dot{M}+(A+BK)M+M(A+BK)^\top-2\lambda M\bigg)\delta_x\leq0
	\end{equation*}
	holds, for every $(x,\delta_x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$. Consequently, the condition defined by the set of equations \eqref{eq:Arstein-Sontag} holds. Thus, $M$ is a sum-separable control-contraction metric for system \eqref{eq:general system} and, according to Theorem~\ref{prop:CCM Existence}, \eqref{eq:differential feedback law} is a differential feedback rendering the equilibrium of the origin globally exponentially stable for system~\eqref{eq:general system:differential} in closed loop.

@@

\mypara 	From the fundamental theorem of calculus, Equation \eqref{eq:contracting feedback law} is equivalent to the differential equation
	\begin{equation}\label{eq:K ODE}
		\tfrac{dk^\ast}{ds}(t,s)=K(\gamma(t,s),k^\ast(t,s))\gamma_s(t,s)=:\overline{K}(t,s,k^\ast)\;.
	\end{equation}
which has a unique solution defined on a maximal interval of existence, for every initial condition, due to the regularity of the functions $K$ and $\gamma$. It remains to show that this interval of existence can be extended on $[0,1]$. In other words, show that there exist no $\bar{s}\in[0,1]$ such that $k^\ast(t,s)$ tends to infinity, as $s$ tends to $\bar{s}$.

@@

\mypara Although Theorem~\ref{thm:main result} provides a methodology to design distributed controllers, \emph{a priori}, nothing has been imposed to compute the MI~\eqref{eq:MI formulation} in a distributed manner. The next section provides sufficient conditions to parallelize the search for a solution to the MI~ \eqref{eq:MI formulation}.

@@

\mypara For large scale systems, the dimension of the matrices $W$ and $Y$ can be a problem for the computation of a solution to the MI~\eqref{eq:MI formulation}. This is due to the fact that semidefinite programming become less efficient when those matrices have high rank \cite{VandenbergheBoyd1996}. For this reason, the computation of a solution to MI~\eqref{eq:MI formulation} need to be done in parallel.

@@

\begin{proposition}\label{prop:clique tree decomposition}
	Let $p\in\mathbb{N}_{>0}$ be the number of nodes of the clique tree $\mathscr{T}(\mathscr{G})$. Then, the MI~\eqref{eq:MI formulation} can be decomposed into $p$ matrix inequalities with smaller dimension. Moreover, if each matrix is negative definite, then the MI~\eqref{eq:MI formulation} is satisfied.
\end{proposition}

@@

\mypara    Denote the left-hand side of the MI~\eqref{eq:MI formulation} by $T$. The existence of $p$ cliques implies that there exist matrices $F_k:\mathbb{R}^{\dim_k}\to\mathbb{R}^{\dim_k\times \dim_k}$, where $k\in\mathbb{N}_{[1,p]}$, satisfying the equation
   \begin{equation*}
   	 T=\sum_{k=1}^p E_k^\top F_k E_k\;.
  	\end{equation*}

\mypara Consequently, whenever the following matrix inequality holds 
  	\begin{equation}\label{eq:MIF}
  		F_k\prec0,\quad\forall k\in\mathbb{N}_{[1,p]},
	\end{equation}
	uniformly over $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$, the matrix $T$ is uniformly negative definite. Thus, the MI~\eqref{eq:MI formulation} holds. This concludes the proof.
\end{proof}

@@

\mypara When $N\geq6$ and $Y$ is a full-rank matrix with elements described by polynomials, the MI \eqref{eq:MI formulation} could not be solved on a standard desktop computer. A reason for this issue is the dependence of each element of the matrix $Y$ on the variable $q$. Another reason is the need to verify if the matrix \eqref{eq:MI formulation} is negative definite. For instance, when  polynomials with degree $d$ are employed, the number of monomials to be optimized in the $(i,j)$ position the matrix $T$ is
\begin{equation*}
		\begin{pmatrix}
n+d\\ d
\end{pmatrix}
\end{equation*}
\cite{Waki2006}. Consequently, when no structural constraints are imposed on $Y$, this verification can lead to
\begin{equation*}
n\dfrac{n+1}{2}\begin{pmatrix}
n+d\\ d
\end{pmatrix}
\end{equation*}
monomials to be computed. This motivates the approach proposed in this paper.

@@

\mypara In the first one, no constraints were imposed on the matrix $Y$ (it is defined as a full matrix). More precisely, for each $k\in\mathbb{N}_{[1,4]}$, $\mathscr{K}(k)=\{1,2,3,4\}$, for each $i\in\mathbb{N}_{[1,4]}$, $\mathbin{\mathtt{proj}}_{\mathscr{K}(i)}(\mathbb{R}^8)=\mathbb{R}^8$, $\mathbf{C}(i)=\mathbb{R}^8$. The parser took 98  seconds to solve the MI \eqref{eq:MI formulation}. The matrix $W$ is given by
\begin{equation*}
		W=\mathbin{\mathtt{diag}}(1,1,2,1,2,1,1,1)
\end{equation*}
and the matrix $Y$ has the structure shown in Equation~\eqref{eq:example:Y:general} with elements
\begin{align*}
 \overline{Y}_{ij}(q)=&-2y_i\;,\\
 \underline{Y}_{ij}(q)=&-2-2\sum_{k=1}^4 y_k^2+x_k^2\;.
\end{align*}

@@

\mypara For the remaining cases, the structure of the network composed of systems of the form \eqref{eq:example} has been exploited. More specifically, Proposition~\ref{prop:clique tree decomposition} can be employed, as the graph describing the interconnection of the network is a subclass of chordal graphs. Consequently, the MI~\eqref{eq:MI formulation} can be decomposed into $N-1$ cliques  described by the matrix
\begin{equation*}
	F_k=\begin{bmatrix}
		F_{11}&F_{12}\\
		F_{12}^\top&F_{22}
	\end{bmatrix}\;.
\end{equation*}

@@

\mypara In the last case, the fully decentralized scenario has been considered. The time taken by the parser to solve the MI~\eqref{eq:MI formulation} was 2.6 seconds. In this case, for each index $i\in\mathbb{N}_{[1,4]}$, $\mathscr{K}(i)=\{i\}$, $\mathbin{\mathtt{proj}}_{\mathscr{K}(i)}(\mathbb{R}^8)=\{q_i\}$. Moreover, the sets $\mathbf{C}$ are defined as
\begin{equation*}
	\begin{array}{rcl}
		\mathbf{C}(1)&=&\{\begin{bmatrix}
		1,&1,&0,&0,&0,&0,&0,&0\end{bmatrix}\},\\
		\mathbf{C}(2)&=&\{\begin{bmatrix}
		0,&0,&1,&1,&0,&0,&0,&0\end{bmatrix}\},\\
		\mathbf{C}(3)&=&\{\begin{bmatrix}
		0,&0,&0,&0,&1,&1,&0,&0\end{bmatrix}\},\\
		\mathbf{C}(4)&=&\{\begin{bmatrix}
		0,&0,&0,&0,&0,&0,&1,&1\end{bmatrix}\}\;.
	\end{array} 
\end{equation*}
Solving the MI \eqref{eq:MI formulation} the matrix $W$ 
\begin{equation*}
 W=\mathbin{\mathtt{diag}}\left(2,1,1,1,1,1,1,1\right),\\
\end{equation*}
and the matrix $Y$ with structure given by \eqref{eq:example:Y:neighbor and decentralized} with elements defined, for every $q_i\in\mathbb{R}^2$, by 
\begin{align*}
 \overline{Y}_i(q_i)=-\begin{bmatrix}2y_i,3+2(x_i^2+y_i^2)\end{bmatrix},\quad i\in\mathbb{N}_{[1,4]}\;.
\end{align*}

\mypara Figure \ref{fig:time graph} shows a plot of the time taken by parser to the MI \eqref{eq:MI formulation} for the three cases considered in this topology: unconstrained, ``neighbor'' and fully decentralized.  According to this graph, for $N=1,2$, the time taken by the parser to solve the MI \eqref{eq:MI formulation} has the same order of magnitude for the three cases. However, as the number of systems increases, the unconstrained case takes more time to be solved than the ``neighbor'' case which, in turn, takes more time than the fully decentralized case. The time difference between the two latter cases can be explained by fact that, for the ``neighbor'' case, the matrix $Y$ contains more non-identically zero elements than the fully decentralized case. In addition to this, in the former case, each element is defined as a polynomial of second degree on the variables $q_i$ and $\breve{q}_i$, while for the latter, they are defined on the variable $q_i$ only.

@@

\mypara In the ``unconstrained'' case, the solver took 7.4s to solve the MI~\eqref{eq:MI formulation} using chordal graphs. The matrix $W$ is given by $\mathbin{\mathtt{diag}}(\overline{W},\overline{W},\overline{W},\overline{W})$, where $\overline{W}=\mathbin{\mathtt{diag}}(0.7,1.2)$. The matrix $Y$ is given by
\begin{equation*}
	Y(q)=\begin{bmatrix}
	\overline{Y}(q_1)&0_{2\times1}&0_{2\times1}&0_{2\times1}\\
	0_{2\times1}&\overline{Y}(q_2)&0_{2\times1}&0_{2\times1}\\
	0_{2\times1}&0_{2\times1}&\overline{Y}(q_3)&0_{2\times1}\\
	0_{2\times1}&0_{2\times1}&0_{2\times1}&\overline{Y}(q_4)
	\end{bmatrix}\;,
\end{equation*}
where $\overline{Y}(q_i)=-\begin{bmatrix}2.4y_i,&2.9+2.9(x_i^2+y_i^2)\end{bmatrix}$.

>>>>

\begin{theorem}\label{thm:main result}
	If there exists a solution to the matrix inequalities
	\begin{subequations}\label{eq:main result}
		\begin{equation}\label{eq:MI:SSCCM}
			-\dot{W}+A+WA^\top+BY+(BY)^\top+2\lambda W\prec0
		\end{equation}
		with $W=\mathbin{\mathtt{diag}}(W_1,\ldots,W_N)$,
		\begin{equation}\label{eq:Ws:SSCCM}
			W_1\succ0,\ldots,W_N\succ0
		\end{equation}
		and
		\begin{equation}\label{eq:YinXi}
			Y\in\Xi.
		\end{equation}
	\end{subequations}
	Then, $W$ is a sum-separable control-contraction metric for system \eqref{eq:general system} and there exists a solution to Problem~\ref{problem formulation}. In particular, the graph $\mathscr{G}_c$ describing the communication network among different controllers specifies the set $\Xi$.
\end{theorem}

\begin{proof}[Proof of Theorem \ref{thm:main result}]
	By assumption, the functions $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ and $Y:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{m\times n}$ satisfy the set of equations~\eqref{eq:main result}. Apply the coordinate change $\eta=M\delta$ and define the matrices $W=M^{-1}$ and $K=YW^{-1}$. Since $W$ is diagonal, the structure of $Y$ is preserved and, consequently, $K\in\Xi$.

\mypara	The MI~\eqref{eq:MI:SSCCM} implies that the inequality
	\begin{equation*}
		\delta_x^\top\bigg(\dot{M}+(A+BK)M+M(A+BK)^\top-2\lambda M\bigg)\delta_x\leq0
	\end{equation*}
	holds, for every $(x,\delta_x,u)\in\mathbb{R}^n\times\mathbb{R}^n\times\mathbb{R}^m$. Consequently, the condition defined by the set of equations \eqref{eq:Arstein-Sontag} holds. Thus, $M$ is a sum-separable control-contraction metric for system \eqref{eq:general system} and, according to Theorem~\ref{prop:CCM Existence}, \eqref{eq:differential feedback law} is a differential feedback rendering the equilibrium of the origin globally exponentially stable for system~\eqref{eq:general system:differential} in closed loop.

@@

\mypara 	From the fundamental theorem of calculus, Equation \eqref{eq:contracting feedback law} is equivalent to the differential equation
	\begin{equation}\label{eq:K ODE}
		\tfrac{dk^\ast}{ds}(t,s)=K(\gamma(t,s))k^\ast(t,s)\gamma_s(t,s)=:\overline{K}(t,s,k^\ast)\;.
	\end{equation}
which has a unique solution defined on a maximal interval of existence, for every initial condition, due to the regularity of the functions $K$ and $\gamma$. It remains to show that this interval of existence can be extended on $[0,1]$. In other words, show that there exist no $\bar{s}\in[0,1]$ such that $k^\ast(t,s)$ tends to infinity, as $s$ tends to $\bar{s}$.

@@

\mypara Although Theorem~\ref{thm:main result} provides a methodology to design distributed controllers, \emph{a priori}, nothing has been imposed to compute the set of equations~\eqref{eq:main result} in a distributed manner. The next section provides sufficient conditions to parallelize the search for a solution to the set of equations~\eqref{eq:main result}.

@@

\mypara For large scale systems, the dimension of the matrices $W$ and $Y$ can be a problem to satisfy the set of equations~\eqref{eq:main result}. This is due to the fact that semidefinite programming become less efficient when those matrices have high rank \cite{VandenbergheBoyd1996}. In particular, the computation of a solution to MI~\eqref{eq:MI:SSCCM} need to be done in parallel.

@@

\begin{proposition}\label{prop:clique tree decomposition}
	Let $p\in\mathbb{N}_{>0}$ be the number of nodes of the clique tree $\mathscr{T}(\mathscr{G})$. Then, the MI~\eqref{eq:MI:SSCCM} can be decomposed into $p$ matrix inequalities with smaller dimension. Moreover, if each matrix is negative definite, then the MI~\eqref{eq:MI:SSCCM} is satisfied.
\end{proposition}

@@

\mypara    Denote the left-hand side of the MI~\eqref{eq:MI:SSCCM} by $T$. The existence of $p$ cliques implies that there exist matrices $F_k:\mathbb{R}^{\dim_k}\to\mathbb{R}^{\dim_k\times \dim_k}$, where $k\in\mathbb{N}_{[1,p]}$, satisfying the equation
   \begin{equation*}
   	 T=\sum_{k=1}^p E_k^\top F_k E_k\;.
  	\end{equation*}

\mypara Consequently, whenever the following matrix inequality holds 
  	\begin{equation}\label{eq:MIF}
  		F_k\prec0,\quad\forall k\in\mathbb{N}_{[1,p]},
	\end{equation}
	uniformly over $(x,u)\in\mathbb{R}^n\times\mathbb{R}^m$, the matrix $T$ is uniformly negative definite. Thus, the MI~\eqref{eq:MI:SSCCM} holds. This concludes the proof.
\end{proof}

@@

\mypara When $N\geq6$ and $Y$ is a full-rank matrix with elements described by polynomials, the MI~\eqref{eq:MI:SSCCM} could not be solved on a standard desktop computer. A reason for this issue is the dependence of each element of the matrix $Y$ on the variable $q$. Another reason is the need to verify if the matrix inequality~\eqref{eq:MI:SSCCM} is negative definite. For instance, when  polynomials with degree $d$ are employed, the number of monomials to be optimized in the $(i,j)$ position the matrix $T$ is
\begin{equation*}
		\begin{pmatrix}
n+d\\ d
\end{pmatrix}
\end{equation*}
\cite{Waki2006}. Consequently, when no structural constraints are imposed on $Y$, this verification can lead to
\begin{equation*}
n\dfrac{n+1}{2}\begin{pmatrix}
n+d\\ d
\end{pmatrix}
\end{equation*}
monomials to be computed. This motivates the approach proposed in this paper.

@@

\mypara In the first one, no constraints were imposed on the matrix $Y$ (it is defined as a full matrix). More precisely, for each $k\in\mathbb{N}_{[1,4]}$, $\mathscr{K}(k)=\{1,2,3,4\}$, for each $i\in\mathbb{N}_{[1,4]}$, $\mathbin{\mathtt{proj}}_{\mathscr{K}(i)}(\mathbb{R}^8)=\mathbb{R}^8$, $\mathbf{C}(i)=\mathbb{R}^8$. The parser took 98  seconds to solve the MI~\eqref{eq:MI:SSCCM}. The matrix $W$ is given by
\begin{equation*}
		W=\mathbin{\mathtt{diag}}(1,1,2,1,2,1,1,1)
\end{equation*}
and the matrix $Y$ has the structure shown in Equation~\eqref{eq:example:Y:general} with elements
\begin{align*}
 \overline{Y}_{ij}(q)=&-2y_i\;,\\
 \underline{Y}_{ij}(q)=&-2-2\sum_{k=1}^4 y_k^2+x_k^2\;.
\end{align*}

@@

\mypara For the remaining cases, the structure of the network composed of systems of the form \eqref{eq:example} has been exploited. More specifically, Proposition~\ref{prop:clique tree decomposition} can be employed, as the graph describing the interconnection of the network is a subclass of chordal graphs. Consequently, the MI~\eqref{eq:MI:SSCCM} can be decomposed into $N-1$ cliques  described by the matrix
\begin{equation*}
	F_k=\begin{bmatrix}
		F_{11}&F_{12}\\
		F_{12}^\top&F_{22}
	\end{bmatrix}\;.
\end{equation*}

@@

\mypara In the last case, the fully decentralized scenario has been considered. The time taken by the parser to solve the MI~\eqref{eq:MI:SSCCM} was 2.6 seconds. In this case, for each index $i\in\mathbb{N}_{[1,4]}$, $\mathscr{K}(i)=\{i\}$, $\mathbin{\mathtt{proj}}_{\mathscr{K}(i)}(\mathbb{R}^8)=\{q_i\}$. Moreover, the sets $\mathbf{C}$ are defined as
\begin{equation*}
	\begin{array}{rcl}
		\mathbf{C}(1)&=&\{\begin{bmatrix}
		1,&1,&0,&0,&0,&0,&0,&0\end{bmatrix}\},\\
		\mathbf{C}(2)&=&\{\begin{bmatrix}
		0,&0,&1,&1,&0,&0,&0,&0\end{bmatrix}\},\\
		\mathbf{C}(3)&=&\{\begin{bmatrix}
		0,&0,&0,&0,&1,&1,&0,&0\end{bmatrix}\},\\
		\mathbf{C}(4)&=&\{\begin{bmatrix}
		0,&0,&0,&0,&0,&0,&1,&1\end{bmatrix}\}\;.
	\end{array} 
\end{equation*}
Solving the MI \eqref{eq:MI formulation} the matrix $W$ 
\begin{equation*}
 W=\mathbin{\mathtt{diag}}\left(2,1,1,1,1,1,1,1\right),\\
\end{equation*}
and the matrix $Y$ with structure given by \eqref{eq:example:Y:neighbor and decentralized} with elements defined, for every $q_i\in\mathbb{R}^2$, by 
\begin{align*}
 \overline{Y}_i(q_i)=-\begin{bmatrix}2y_i,3+2(x_i^2+y_i^2)\end{bmatrix},\quad i\in\mathbb{N}_{[1,4]}\;.
\end{align*}

\mypara Figure \ref{fig:time graph} shows a plot of the time taken by parser to the MI \eqref{eq:MI:SSCCM} for the three cases considered in this topology: unconstrained, ``neighbor'' and fully decentralized.  According to this graph, for $N=1,2$, the time taken by the parser to solve the MI~\eqref{eq:MI:SSCCM} has the same order of magnitude for the three cases. However, as the number of systems increases, the unconstrained case takes more time to be solved than the ``neighbor'' case which, in turn, takes more time than the fully decentralized case. The time difference between the two latter cases can be explained by fact that, for the ``neighbor'' case, the matrix $Y$ contains more non-identically zero elements than the fully decentralized case. In addition to this, in the former case, each element is defined as a polynomial of second degree on the variables $q_i$ and $\breve{q}_i$, while for the latter, they are defined on the variable $q_i$ only.

@@

\mypara In the ``unconstrained'' case, the solver took 7.4s to solve the MI~\eqref{eq:MI:SSCCM} using chordal graphs. The matrix $W$ is given by $\mathbin{\mathtt{diag}}(\overline{W},\overline{W},\overline{W},\overline{W})$, where $\overline{W}=\mathbin{\mathtt{diag}}(0.7,1.2)$. The matrix $Y$ is given by
\begin{equation*}
	Y(q)=\begin{bmatrix}
	\overline{Y}(q_1)&0_{2\times1}&0_{2\times1}&0_{2\times1}\\
	0_{2\times1}&\overline{Y}(q_2)&0_{2\times1}&0_{2\times1}\\
	0_{2\times1}&0_{2\times1}&\overline{Y}(q_3)&0_{2\times1}\\
	0_{2\times1}&0_{2\times1}&0_{2\times1}&\overline{Y}(q_4)
	\end{bmatrix}\;,
\end{equation*}
where $\overline{Y}(q_i)=-\begin{bmatrix}2.4y_i,&2.9+2.9(x_i^2+y_i^2)\end{bmatrix}$.
